---
title: "Reproducible Research - Week 1 Project"
author: "John Koegel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment
1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## 1. Code for reading in the dataset and/or processing the data

```{r}
# Check Installed Packages, Install Necessary Packages, and Load Packages
list.of.packages <- c("dplyr",'ggplot2','gridExtra','lubridate')
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages,library, character.only = TRUE)
rm('list.of.packages','new.packages')

# Set Working Directory
setwd("~/Google Drive/Development/R/Coursera/Hopkins Data Science Specialization/Reproducible Research/Week 1 Project")
dataDir <- paste(getwd(),"/data/",sep = "")
# Download Data
if (!file.exists(".data")) {
        dir.create(dataDir)
}

rawDataURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists(paste(dataDir,"/data.zip",sep=""))) {
        download.file(rawDataURL, destfile = paste(dataDir,"/data.zip",sep=""), method= "auto")
        unzip(zipfile=paste(dataDir,"/data.zip",sep=""),exdir = "data")
}

# Load activity data into table
activity <- read.csv(paste(dataDir,"activity.csv", sep=""), header = TRUE)
activityDfRAW <- tbl_df(activity)
rm(activity)
activityDf <- filter(activityDfRAW, !is.na(steps))

activityDaily <- group_by(activityDf, date) %>% summarise(steps = sum(steps))
```

## 2. Histogram of the total number of steps taken each day

```{r}
hist(activityDaily$steps)
```

## 3. Mean and median number of steps taken each day

```{r}
activityDailySummary <- summarise(activityDaily,mean(steps), median(steps))
activityDailySummary
```
## 4. Time series plot of the average number of steps taken
```{r}
activityDailyTS <- group_by(activityDf, date) %>% summarise(steps = mean(steps))
ggplot(activityDailyTS) + geom_point(aes(activityDailyTS$date,activityDailyTS$steps))
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
activity5Min <- group_by(activityDf, interval) %>% summarise(steps = mean(steps))
filter(activity5Min, steps == max(activity5Min$steps))
```

## 6. Code to describe and show a strategy for imputing missing data
Missing values should be filled in with the average values for that timeframe.
```{r}
activityDfNA <- filter(activityDfRAW, is.na(steps))
activityDfNA <- full_join(activityDfNA,activity5Min, by = c("interval" = "interval")) %>% select(-steps.x) %>% mutate(steps = steps.y) %>% select(-steps.y)
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed
Imputed values merged back with clean data set, grouped, and then a histogram is created
```{r}
full <- rbind(activityDfNA, activityDf)
fullActivityDaily <- group_by(full, date) %>% summarise(steps = sum(steps))
hist(fullActivityDaily$steps)
```
## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
# Determine Weekdays
fullActivity5Min <- mutate(full, wDay = wday(date))
weekendActivity5Min <- fullActivity5Min %>% filter(wDay ==1 | wDay == 7) %>% select(-wDay) %>% group_by(interval) %>% summarise(steps = mean(steps))
wdayActivity5Min <- fullActivity5Min %>% filter(wDay != 1 & wDay != 7) %>% select(-wDay) %>% group_by(interval) %>% summarise(steps = mean(steps))

myplot1<-ggplot(weekendActivity5Min, aes(interval, steps))+geom_point(color="firebrick") + ggtitle('Weekend Activity')
myplot2<-ggplot(wdayActivity5Min, aes(interval, steps))+geom_point(color="olivedrab") + ggtitle('Week Activity')
grid.arrange(myplot1, myplot2, ncol=2)
```

